The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-01 paper.tex
small changes section 2

\section{What does Anthropomorphism mean?}
\label{sec:anthropomorphism}

Anthropomorphism is an interesting phenomenon because it tells us something about 
the relationship between the human who anthropomorphizes, and the non-human agent 
which is anthropomorphized. Though commonly observed, it still appears controversial, at 
the same time, that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}. 
To shed light on this phenomenon, first of all, we would like to propose a clarification of terms.
In short, we propose to differentiate between \textbf{anthropomorphic form} to refer to an imitation of human-like form in the \textit{design} of artificial agents, and  \textbf{anthropomorphism} to refer to the social phenomenon of 
people's tendency to \textit{perceive} human-likeness in non human-like agents
\citep{fink_anthropomorphism_2012}. In the following, we might use the terms \texit{form} and \textit{design} interchangeably.

\hl{Julia: I will make a draft of a figure here}

As anthropomorphism arises in an interaction between a human and a non-human agent
\citep{persson_anthropomorphism_2000}, one can also
take two main perspectives on explaining anthropomorphism \citep{lee_human_2005}: an artifact-centered perspective and a human-centered perspective.\\ 
We now first present an artifact-centered perspective, and outline how 
robotics uses \textit{anthropomorphic design} to encourage 
people to interact in a social, human-like way with robots.
Then, we present a human-centered 
perspective and shed light on \textit{anthropomorphism as a social phenomenon}. We present 
theories from developmental psychology, cognitive science, and social psychology to explain people's tendency to 
perceive human-like characteristics in non-human agents.
It has to be kept in mind that both perspectives are inseparable; they apply at the 
same time, and we would like to present them here just as two different viewpoints 
on the same phenomenon.
In the end of this section, we synthesize 
the presented backgrounds into an extended understanding of anthropomorphism 
as a multi-factor phenomenon.

@@

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize it. 
This perspective assumes that humans directly (mindlessly) respond to life-like and social 
cues that the non-human agent or system emits \citep{nass_machines_2000}. Without thoughtful mental processing, 
humans tend to simply apply stereotypes and heuristics to the non-human agent, and in turn apply human-human social schemas and norms to the occurring interactions.
\cite{takayama_perspectives_2012} propose to refer to this immediate (sometimes visceral) sense in a situation as \textit{in-the-moment}, compared to a more distanced cogitation and consideration which she calls \textit{reflective}. (As this distinction imply different cognitive processes, we will come back to it later, in the human-centered perspective, Section \ref{sec:human-centered}).

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
Apart from this, it is suggested that an increased human-likeness of a robot (up to a 
certain threshold) will generally lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \citep{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
that humans tend to project everyday 
expectations of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \citep{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus, the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot. 
These experiences mark the transition between an immediate \textit{in-the-moment} and more thoughtful \textit{reflective} \citep{takayama_perspectives_2012} perception of, and interaction with the robot. As such, a more human-centered perspective needs to be applied, and we will come back to this in Section \ref{sec:human-centered}.

@@

This cognitive viewpoint on anthropomorphism is important because people's estimation of an agent's knowledge model and its capabilities affects the way they relate to it. This holds important implications for the resulting interaction with the system, the user experience, and the acceptance. Previous research examined the validity of the mental model concept with various kinds of robots \citep{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people tend to hold richer mental models of anthropomorphic robots in contrast to mechanic ones \citep{kiesler_mental_2002}.\footnote{\hl{Later in this article, we may however come to question how far this result is based on expectations from the human user side, which are likely to be refined after continued interaction, and after the user gets acquainted with the robot's behavior.}} 
A similar finding is described in \cite{hegel_understanding_2008} and
\citet{krach_can_2008}. In a user study using functional magnetic resonnance
imagery (fMRI), the authors found that participants
implicitly attributed human-like qualities, such as mental states, to their
non-human interaction partner. Indeed, from an early age humans develop a
tendency to explain one's own and others'actions in terms of beliefs, desires
and goals,  called \textit{theory of mind}. The theory of mind allows the implicit attribution of
intentions and other mental states to others \citep{premack1978does, leslie_pretense_1987, Frith2003}.
\citet{hegel_understanding_2008} and \citet{krach_can_2008} finding was evident at the behavioral level
as well as on the neuro-physiological level: the more the human-likeness of the
artefactual partner increases, the more brain areas associated with theory of
mind  get activated
\citep{krach_can_2008}. \footnote{\hl{Ha! This says that anthropomorphism can also occur on a behavioral level, yes! :) so we should also measure anthropomorphism in the behavior.}} 
Implication of human brain area involved in the inference of others` mental
states has also been shown in response to viewing
non-anthropomorphic/non-humanoid robotic gadgets whose behavior
has been described as unpredictable -- but not in response to those whose
behavior was described as predictable -- a finding also found at the behavioral
level \citep{Waytz2010}.
Interestingly, the grey matter volume of a brain area related to theory of mind
has been correlated to individual's score of anthropomorphism
\citep{cullen2013individual}.

@@

\subsubsection{Diverse understandings of anthropomorphism\\}

@@

>>>>

\section{Why do humans anthropomorphize?}
\label{sec:anthropomorphism}

Anthropomorphism is an interesting phenomenon because it tells us something about 
the relationship between the human who anthropomorphizes, and the non-human agent 
which is anthropomorphized. Though commonly observed, it still appears controversial, at 
the same time, that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}. 
To shed light on this phenomenon as a whole, first of all, we would like to propose a clarification of terms.

\paragraph{Terminology\\}

We propose to use the term \textbf{anthropomorphism} to describe a social phenomenon, which arises in an interaction between a human agent and a non-human agent, \eg a robot (see Figure \ref{fig:anthropofig}). The term anthropomorphism can be used to refer to people's tendency to \textit{perceive} human-like characteristics in non-human agents (or artifacts), but also to refer to the observable effect of a person \textit{ascribing} or \textit{attributing} such human-like characteristics to the agent or artifact. Further, anthropomorphism might also be observable in a human's behavior / actions toward a non-human agent. The common point here is that anthropomorphism is always performed by the human (either in real or imagined).

However, the term anthropomorphism should \textit{not} be used to refer to the design of a non-human agent or artifact. This would violate the understanding of anthropomorphism arising in an interaction. We propose to use the term  \textbf{anthropomorphic design} to refer to an imitation of human-like form in the \textit{form} of artificial agents. (In the following, we might use the terms \textit{form} and \textit{design} interchangeably.)

In literature, this distinction is not always done, which is a major source of confusion.\\

\begin{figure}
    \centering
    \includegraphics[width=0.75\columnwidth]{anthropo.pdf}
    \caption{Anthropomorphism is a social phenomenon which arises in an interaction between a human agent and a non-human agent. The two agents are the main \textit{players} \hl{(or elements?)} of the interaction, and consequently one can apply two main perspectives when trying to explain anthropomorphism: a human-centered and an artifact-centered perspective (grey triangles). There are several person- and artifact-related factors that determine \hl{(or influence / impact on?)} anthropomorphism. Further, the context / situation in which the interaction occurs plays a role.}

    \label{fig:anthropofig}
\end{figure}

As anthropomorphism arises in an interaction between a human and a non-human agent
\citep{persson_anthropomorphism_2000}, one can also
take two main \textit{perspectives} when trying to explain why humans anthropomorphize \citep{lee_human_2005}: an artifact-centered perspective and a human-centered perspective (see grey triangles in Figure \ref{fig:anthropofig}).

We now first present an artifact-centered perspective, and outline how 
robotics uses \textit{anthropomorphic design} to encourage 
people to interact in a social, human-like way with robots.
Then, we present a human-centered 
perspective and shed light on \textit{anthropomorphism as a social phenomenon}. We present 
theories from developmental psychology, cognitive science, and social psychology to explain people's tendency to 
perceive human-like characteristics in non-human agents.
It has to be kept in mind that both perspectives are inseparable; they apply at the 
same time, and we would like to present them here just as two different viewpoints 
on the same phenomenon.
In the end of this section, we synthesize 
the presented backgrounds into an extended understanding of anthropomorphism 
as a multi-factor phenomenon.

@@

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize it. 
This perspective assumes that humans directly (mindlessly) respond to life-like and social 
cues that the non-human agent or system emits \citep{nass_machines_2000}. Without thoughtful mental processing, 
humans tend to simply apply stereotypes and heuristics to the non-human agent, and in turn apply human-human social schemas and norms to the occurring interactions.
\cite{takayama_perspectives_2012} propose to refer to this immediate (sometimes visceral) sense in a situation as \textit{in-the-moment}, compared to a more distanced cogitation and consideration which she calls \textit{reflective}. (As this distinction implies different cognitive processes, we will come back to it later, in the human-centered perspective, Section \ref{sec:human-centered}).

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
Apart from this, it is suggested that an increased human-likeness of a robot (up to a 
certain threshold) will generally lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970}.
% (see Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \citep{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
that humans tend to project everyday 
expectations of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \citep{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus, the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot. 
These experiences mark the transition between an immediate \textit{in-the-moment} and more thoughtful \textit{reflective} \citep{takayama_perspectives_2012} perception of, and interaction with the robot. As this distinction is also based on different cognitive processes, a more human-centered / cognitive perspective needs to be applied here, and we will treat this in Section \ref{sec:human-centered}.

@@

This cognitive viewpoint on anthropomorphism is important because people's estimation of an agent's knowledge model and its capabilities affects the way they relate to it. This holds important implications for the resulting interaction with the system, the user experience, and the acceptance. Previous research examined the validity of the mental model concept with various kinds of robots \citep{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people tend to hold richer mental models of anthropomorphic robots in contrast to mechanic ones \citep{kiesler_mental_2002}.\footnote{\hl{Later in this article, we may however come to question how far this result is based on expectations from the human user side, which are likely to be refined after continued interaction, and after the user gets acquainted with the robot's behavior.}} 
A similar finding is described in \cite{hegel_understanding_2008} and
\cite{krach_can_2008}. In a user study using functional magnetic resonance
imagery (fMRI), the authors found that participants
implicitly attributed human-like qualities, such as mental states, to their
non-human interaction partner. Indeed, from an early age humans develop a
tendency to explain one's own and others' actions in terms of beliefs, desires
and goals,  called \textit{theory of mind}. The theory of mind allows the implicit attribution of
intentions and other mental states to others \citep{premack1978does,leslie_pretense_1987,Frith2003}.
The finding of \cite{hegel_understanding_2008} and \cite{krach_can_2008} was evident at the behavioral level
as well as on the neuro-physiological level: the more the human-likeness of the
artefactual partner increases, the more brain areas associated with theory of
mind  get activated
\citep{krach_can_2008}. \footnote{\hl{Ha! This says that anthropomorphism can also occur on a behavioral level, yes! :) so we should also measure anthropomorphism in the behavior.}} 
Implication of human brain area involved in the inference of others` mental
states has also been shown in response to viewing
non-anthropomorphic/non-humanoid robotic gadgets whose behavior
has been described as unpredictable -- but not in response to those whose
behavior was described as predictable -- a finding also found at the behavioral
level \citep{Waytz2010}.
Interestingly, the grey matter volume of a brain area related to theory of mind
has been correlated to individual's score of anthropomorphism
\citep{cullen2013individual}.

@@

\subsubsection{Diverse understandings of anthropomorphism and confusions in the terminology used\\}

@@

\hl{Julia is working on this section}
